
\null\vfill
\begin{center}
\textit{To the Most Holy Trinity, Undivided Unity.}
\end{center}
\vfill

\pagebreak

\thispagestyle{plain}
\null\vfill

\section*{Thesis Statement}
Advancing science and engineering with deep learning-based emergent language requires goal-oriented, environment-agnostic analytical algorithms.

\section*{Abstract}
Emergent communication is the field of research which studies how human language-like communication systems evolve from scratch in agent-based simulations.
Such simulations have been studied since the 1990s, yet in 2016, researchers introduced recent advancements in deep neural networks, reinforcement learning, and natural language processing into the field.
Deep learning-based emergent communication, as a paradigm, has significant potential applications from powering unprecedentedly detailed simulations of how humans invent, acquire, and use language
  to providing rich, embodied language data for training and evaluating models free of contamination or surveillance.
Despite this potential, the field has yet to make progress towards the more revolutionary applications because it lacks the methodological resources to enable cumulative research.
The design space of emergent language is dizzyingly large, and the outputs of said environments are similarly difficult to interpret.
As a result, analyses of emergent languages, and consequently their findings, are often ``one-off'', unable to build on prior work or produce generalizable contributions.

This thesis, then, advances the field of emergent communication by developing the resources necessary for a cumulative research paradigm, something that is critical to the advancement of any area of science or engineering.
Specifically, it introduces analytical methods for deep learning-based emergent communication that enable measurable progress in the field with a particular mind towards solving practical applications and improving scientific understanding.
I first review the particular ways in which emergent communication can solve practical applications and improve scientific understanding.
This is followed by introducing emergent language data resources which enable empirical evaluation across a variety of emergent languages.
These resources are then used to develop
  (1) a deep transfer learning-based evaluation metric for emergent communication to measure the practical applicability of emergent language
  and (2) algorithms for discovering the morphology of emergent languages as a foundation for further linguistic analysis.

\vfill

\pagebreak
\thispagestyle{plain}

\null\vfill
\section*{Acknowledgments}

I would like to thank first and foremost my adviser, David Mortensen, whose unique interests and expertise has made this unique thesis possible.
The rest of my committee: Yonatan Bisk, Daniel Fried, and Kenny Smith, have joined him setting a standard for my work befitting so great a department, school, and university.
A special thanks goes to Jean Oh who advised me for my first year at Carnegie Mellon University, getting me on my feet in my graduate education.
Gratitude is also due to the peers whom I have worked alongside:
    Charis Graham,
    Shrimai Prabhumoye,
    Chan Young Park,
    Felix Labelle,
    and Artidoro Pagnoni
  as well as to the labs and organizations that have hosted me:
  ChangeLing Lab,
    Linguistics Lab,
    BIG Lab,
    and EvenUp.
Finally, on the home front, I thank my
  mother and father: Yvonne and Brian;
  wife and children: Elizabeth and Siena, Avila, Lisieux, and Magnus;
  and my neighbors in Brookline who have helped make Pittsburgh a home.

\vfill

